RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-05-13 paper/RDF validation requirements/RDF validation requirements.tex
disco validation rules
CWA UNA dependency

@@

\section{Existential Quantification on Properties}

@@

\section{Universal Quantification on Properties}

@@

\section{Property Domain}

@@

\section{Property Range}

@@

\section{Minimum Unqualified Cardinality Restrictions on Properties}

@@

\section{Minimum Qualified Cardinality Restrictions on Properties}

@@

\section{Maximum Unqualified Cardinality Restrictions on Properties}

@@

\section{Maximum Qualified Cardinality Restrictions on Properties}

@@

\section{Exact Unqualified Cardinality Restrictions on Properties}

@@

\section{Exact Qualified Cardinality Restrictions on Properties}

@@

@@

\section{Define Order}

@@

\section{Default Values}

@@

\section{Exploiting Class/Property Specialization Ontology Axioms}

@@

@@

\begin{evaluation-overall}
Context-Specific Valid Classes & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Context-Specific Valid Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Default Values & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ & \ding{55} & $\checkmark$ \\
Mathematical Operations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Matching & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Cardinality & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Whitespace Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
HTML Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Conditional Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Recommended Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Handle RDF Collections & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Value is Valid for Datatype & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Exploiting Class/Property Specialization Ontology Axioms & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
*Cardinality Shortcuts & \ding{55} & $\checkmark$ & \ding{55} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
Aggregations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
\end{evaluation-overall}

@@

\begin{evaluation-generic}
Context-Specific Valid Classes & \ding{55} & $\checkmark$ & \ding{55} \\   
Context-Specific Valid Properties & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Default Values & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Mathematical Operations & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Language Tag Matching & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Language Tag Cardinality & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Whitespace Handling & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
HTML Handling & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Conditional Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Recommended Properties & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Handle RDF Collections & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Value is Valid for Datatype & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
*Exploiting Class/Property Specialization Ontology Axioms & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
*Cardinality Shortcuts & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Aggregations & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
\end{evaluation-generic}

@@

>>>>

\newenvironment{evaluation-cwa-una-dependency}{
  \scriptsize
  \sffamily
  \vspace{0cm}
	\begin{center}
  \begin{tabular}{l|c}
  \hline
  \textbf{Constraint Type} & \textbf{CWA/UNA Dependent} \\
  \hline

\newenvironment{cwa-una-dependency}{
  \scriptsize
  \sffamily
  \vspace{0cm}
	\begin{center}
  \begin{tabular}{l|c|c}
  \hline
  \textbf{Constraint Type} & \textbf{CWA/UNA Dependent} & \textbf{CWA/UNA Dependent} \\
  \hline

@@

\section{Existential Quantifications}

@@

\section{Universal Quantifications}

@@

\section{Property Domains}

@@

\section{Property Ranges}

@@

\section{Minimum Unqualified Cardinality Restrictions}

@@

\section{Minimum Qualified Cardinality Restrictions}

@@

\section{Maximum Unqualified Cardinality Restrictions}

@@

\section{Maximum Qualified Cardinality Restrictions}

@@

\section{Exact Unqualified Cardinality Restrictions}

@@

\section{Exact Qualified Cardinality Restrictions}

@@

\subsection{Class-Specific Irreflexive Object Properties*}

A property is \emph{irreflexive} if it is never locally reflexive \cite{Kroetzsch2012}.
An object property irreflexivity axiom states that the object property \emph{OP} is irreflexive - that is, no individual is connected by \emph{OP} to itself.
\emph{Class-Specific Irreflexive Object Properties} are object properties which are irreflexive within a given context, e.g. a class.

@@

\section{Ordering}

@@

\section{Default Values*}

@@

\section{Use Sub-Super Relations in Validation}

@@

\section{Provenance}

\section{Data Model Consistency}

Is the data consistent with the intended semantics of the data model?
Such validation rules ensure the integrity of the data according to the data model.

\section{Structure}

SKOS is based on RDF, which is a graph-based data model. Therefore we can concentrate on the vocabulary's graph-based structure for assessing the quality of SKOS vocabularies and apply graph- and network-analysis techniques.

\section{Labeling and Documentation}

\section{Vocabulary}
Vocabularies should not invent any new terms or use deprecated elements.

\section{HTTP URI Scheme Violation}

@@

\begin{evaluation-overall}
Context-Specific Valid Classes & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Context-Specific Valid Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Default Values* & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ & \ding{55} & $\checkmark$ \\
Mathematical Operations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Matching & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Cardinality & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Whitespace Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
HTML Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Conditional Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Recommended Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Handle RDF Collections & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Value is Valid for Datatype & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Exploiting Class/Property Specialization Ontology Axioms & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
*Cardinality Shortcuts & \ding{55} & $\checkmark$ & \ding{55} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
Aggregations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
\end{evaluation-overall}

@@

\begin{evaluation-generic}
Context-Specific Valid Classes & \ding{55} & $\checkmark$ & \ding{55} \\   
Context-Specific Valid Properties & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Default Values* & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Mathematical Operations & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Language Tag Matching & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Language Tag Cardinality & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Whitespace Handling & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
HTML Handling & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Conditional Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Recommended Properties & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Handle RDF Collections & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Value is Valid for Datatype & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
*Exploiting Class/Property Specialization Ontology Axioms & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
*Cardinality Shortcuts & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Aggregations & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
\end{evaluation-generic}

@@

\subsection{CWA and UNA Dependency}

\begin{itemize}
	\item Dependent on the \emph{CWA}: do further triples change the validation result?
	\item Dependent on the \emph{UNA}: Do validation results changes in case two resources are the same?
\end{itemize}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcc@{}}
           & \multicolumn{2}{c}{\textbf{Dependency}}
    \\  \cmidrule{2-3}
    \\       \textbf{Constraint Types}
           & \rot{\emph{CWA}}
           & \rot{\emph{UNA}}
    \\ \midrule
		*Subsumption & $\checkmark$ & $\checkmark$ \\
		*Class Equivalence & $\checkmark$ & $\checkmark$ \\
		*Sub Properties & $\checkmark$  & $\checkmark$ \\
		*Property Domain & $\checkmark$ & $\checkmark$ \\
		*Property Range & $\checkmark$ & $\checkmark$ \\
		*Inverse Object Properties & $\checkmark$ & $\checkmark$ \\
		*Symmetric Object Properties & $\checkmark$ & $\checkmark$ \\
		*Asymmetric Object Properties & \ding{55} & \ding{55} \\
		*Reflexive Object Properties & $\checkmark$ & $\checkmark$ \\
		*Irreflexive Object Properties & \ding{55} & \ding{55} \\
		Disjoint Properties & \ding{55} & \ding{55} \\
		Disjoint Classes & \ding{55} & \ding{55} \\
		Context-Sp. Property Groups & $\checkmark$ & $\checkmark$ \\
		Context-Sp. Inclusive OR of P. & $\checkmark$ & $\checkmark$ \\
		Context-Sp. Inclusive OR of P. Groups & $\checkmark$ & $\checkmark$ \\
		Recursive Queries & \ding{55} & \ding{55} \\
		Individual Inequality & \ding{55} & \ding{55} \\
		*Equivalent Properties & $\checkmark$ & $\checkmark$ \\
		Property Assertions & $\checkmark$ & $\checkmark$ \\
		Data Property Facets & \ding{55} & \ding{55} \\
		Literal Pattern Matching & \ding{55} & \ding{55} \\
		Negative Literal Pattern Matching & \ding{55} & \ding{55} \\
		*Object Property Paths & $\checkmark$ & $\checkmark$ \\
		*Intersection & $\checkmark$ & $\checkmark$ \\
		*Disjunction & $\checkmark$ & $\checkmark$ \\
		*Negation & \ding{55} & $\checkmark$ \\
		*Existential Quantification & $\checkmark$ & \ding{55} \\
		*Universal Quantification & \ding{55} & $\checkmark$ \\
		*Minimum Unqualified Cardinality & $\checkmark$ & $\checkmark$ \\
		*Minimum Qualified Cardinality & $\checkmark$ & $\checkmark$ \\
		*Maximum Unqualified Cardinality & \ding{55} & $\checkmark$ \\
		*Maximum Qualified Cardinality & \ding{55} & $\checkmark$ \\
		*Exact Unqualified Cardinality & $\checkmark$ & $\checkmark$ \\
		*Exact Qualified Cardinality & $\checkmark$ & $\checkmark$ \\
		*Transitive Object Properties & $\checkmark$ & $\checkmark$ \\
		Context-Sp. Exclusive OR of P. & \ding{55} & \ding{55} \\
		Context-Sp. Exclusive OR of P. Groups & \ding{55} & \ding{55} \\	
		Allowed Values & \ding{55} & \ding{55} \\
		Not Allowed Values & \ding{55} & \ding{55} \\
		Literal Ranges & \ding{55} & \ding{55} \\
		Negative Literal Ranges & \ding{55} & \ding{55} \\
		Required Properties & $\checkmark$ & $\checkmark$ \\
		Optional Properties & \ding{55} & \ding{55} \\
		Repeatable Properties & $\checkmark$ & \ding{55} \\
		Negative Property Constraints & \ding{55} & $\checkmark$ \\
		*Individual Equality & $\checkmark$ & \ding{55} \\
		*Functional Properties & $\checkmark$ & $\checkmark$ \\
		*Inverse-Functional Properties & $\checkmark$ & $\checkmark$ \\
		*Value Restrictions & $\checkmark$ & $\checkmark$ \\
		*Self Restrictions & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
		\label{tab:evaluation-cwa-una-dependency-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcc@{}}
           & \multicolumn{2}{c}{\textbf{Dependency}}
    \\  \cmidrule{2-3}
    \\       \textbf{Constraint Types}
           & \rot{\emph{CWA}}
           & \rot{\emph{UNA}}
    \\ \midrule
		Primary Key Properties & $\checkmark$ & $\checkmark$ \\
		*Class-Specific Property Range & $\checkmark$ & \ding{55} \\
		*Class-Sp. Reflexive Object P. & $\checkmark$ & $\checkmark$ \\
		Membership in Controlled Vocabularies & $\checkmark$ & $\checkmark$ \\
		IRI Pattern Matching & \ding{55} & \ding{55} \\
		Literal Value Comparison & \ding{55} & \ding{55} \\
		Define Order & $\checkmark$ & $\checkmark$ \\
		Validation Levels & $\checkmark$ & $\checkmark$ \\
		String Operations & \ding{55} & \ding{55} \\
		Context-Specific Valid Classes & \ding{55} & $\checkmark$ \\
		Context-Specific Valid Properties & \ding{55} & $\checkmark$ \\
		Default Values* & $\checkmark$ & $\checkmark$ \\
		Mathematical Operations & \ding{55} & \ding{55} \\
		Language Tag Matching & $\checkmark$ & $\checkmark$ \\
		Language Tag Cardinality & $\checkmark$ & $\checkmark$ \\
		Whitespace Handling & \ding{55} & \ding{55} \\
		HTML Handling & \ding{55} & \ding{55} \\
		Conditional Properties & $\checkmark$ & $\checkmark$ \\
		Recommended Properties & $\checkmark$ & $\checkmark$ \\
		Handle RDF Collections & \ding{55} & \ding{55} \\
		Value is Valid for Datatype & \ding{55} & \ding{55} \\
		*Exploiting Class/Property Specialization Ontology Axioms & \ding{55} & $\checkmark$ \\
		*Cardinality Shortcuts & $\checkmark$ & $\checkmark$ \\ 
		Aggregations & \ding{55} & \ding{55} \\

		Class-Specific Irreflexive Object Properties* & \ding{55} & $\checkmark$ \\
		Provenance & $\checkmark$ & $\checkmark$ \\ 
		Data Model Consistency & $\checkmark$ & $\checkmark$ \\ 
		Structure & $\checkmark$ & $\checkmark$ \\  
		Labeling and Documentation & $\checkmark$ & $\checkmark$ \\ 
		Vocabulary & $\checkmark$ & \ding{55} \\
		HTTP URI Scheme Violation & \ding{55} & \ding{55} \\
    \bottomrule
    \end{tabular}
		\label{tab:evaluation-cwa-una-dependency-2}
    \end{center}
\end{table}

%\subsection{Evaluation of CWA/UNA Dependency}
%
%
%
%\begin{evaluation-cwa-una-dependency}
%*Subsumption & $\checkmark$ \\
%*Class Equivalence & \ding{55} \\
%*Sub Properties & $\checkmark$ \\
%*Property Domain & $\checkmark$ \\ 
%*Property Range & $\checkmark$ \\ 
%*Inverse Object Properties & $\checkmark$ \\ 
%*Symmetric Object Properties & $\checkmark$ \\ 
%*Asymmetric Object Properties & $\checkmark$ \\ 
%*Reflexive Object Properties & $\checkmark$ \\ 
%*Irreflexive Object Properties & $\checkmark$ \\
%Disjoint Properties & \ding{55} \\ 
%Disjoint Classes & \ding{55} \\
%Context-Sp. Property Groups & \ding{55} \\ 
%Context-Sp. Inclusive OR of P. & \ding{55} \\
%Context-Sp. Inclusive OR of P. Groups & \ding{55} \\ 
%Recursive Queries & \ding{55} \\ 
%Individual Inequality & $\checkmark$ \\ 
%*Equivalent Properties & \ding{55} \\ 
%Property Assertions & \ding{55} \\ 
%Data Property Facets & \ding{55} \\ 
%Literal Pattern Matching & \ding{55} \\ 
%Negative Literal Pattern Matching & \ding{55} \\ 
%*Object Property Paths & $\checkmark$ \\ 
%*Intersection & $\checkmark$ \\ 
%*Disjunction & $\checkmark$ \\ 
%*Negation & $\checkmark$ \\ 
%*Existential Quantification & $\checkmark$ \\ 
%*Universal Quantification & $\checkmark$ \\ 
%*Minimum Unqualified Cardinality & $\checkmark$ \\ 
%*Minimum Qualified Cardinality & $\checkmark$ \\ 
%*Maximum Unqualified Cardinality & $\checkmark$ \\ 
%*Maximum Qualified Cardinality & $\checkmark$ \\ 
%*Exact Unqualified Cardinality & $\checkmark$ \\ 
%*Exact Qualified Cardinality & $\checkmark$ \\ 
%*Transitive Object Properties & $\checkmark$ \\ 
%Context-Sp. Exclusive OR of P. & $\checkmark$ \\ 
%Context-Sp. Exclusive OR of P. Groups & $\checkmark$ \\
%Allowed Values & $\checkmark$ \\ 
%Not Allowed Values & $\checkmark$ \\ 
%Literal Ranges & \ding{55} \\ 
%Negative Literal Ranges & \ding{55} \\ 
%Required Properties & $\checkmark$ \\ 
%Optional Properties & \ding{55} \\
%Repeatable Properties & $\checkmark$ \\ 
%Negative Property Constraints & $\checkmark$ \\ 
%*Individual Equality & $\checkmark$ \\ 
%*Functional Properties & $\checkmark$ \\
%*Inverse-Functional Properties & $\checkmark$ \\ 
%*Value Restrictions & $\checkmark$ \\
%*Self Restrictions & $\checkmark$ \\
%Primary Key Properties & $\checkmark$ \\ 
%*Class-Specific Property Range & $\checkmark$ \\ 
%*Class-Sp. Reflexive Object P. & $\checkmark$ \\ 
%Membership in Controlled Vocabularies & $\checkmark$ \\ 
%IRI Pattern Matching & \ding{55} \\ 
%Literal Value Comparison & \ding{55} \\ 
%Define Order & \ding{55} \\ 
%Validation Levels & \ding{55} \\ 
%String Operations & \ding{55} \\ 
%\end{evaluation-cwa-una-dependency}
%
%\begin{evaluation-cwa-una-dependency}
%Context-Specific Valid Classes & \ding{55} \\
%Context-Specific Valid Properties & \ding{55} \\ 
%Default Values & \ding{55} \\ 
%Mathematical Operations & \ding{55} \\ 
%Language Tag Matching & \ding{55} \\ 
%Language Tag Cardinality & $\checkmark$ \\ 
%Whitespace Handling & \ding{55} \\ 
%HTML Handling & \ding{55} \\ 
%Conditional Properties & $\checkmark$ \\ 
%Recommended Properties & \ding{55} \\ 
%Handle RDF Collections & \ding{55} \\ 
%Value is Valid for Datatype & \ding{55} \\ 
%*Exploiting Class/Property Specialization Ontology Axioms & \ding{55} \\ 
%*Cardinality Shortcuts & $\checkmark$ \\ 
%Aggregations & \ding{55} \\ 
%\end{evaluation-cwa-una-dependency}

% -----
